Embeddings and Attention in Predictive Modeling

### 2022-04-11 manuscript/manuscript.Rmd
Finalize manuscript

@Huang2020 refines this approach in a model they call the TabTransformer, which implements two main changes to the approach shown in Equations \ref{eq:embed_to_attn_to_nn}. The first of these is that, instead of using the matrix of embeddings after processing with self-attention ($X'$) directly in the model, the output of a Transformer model is used instead. The second of these refinements is that for each of the $q$ covariates with an embedding in the matrix $X'$, we include a new row embedding which denotes which covariate each embedding relates to. Thus, in the TabTransformer model, we augment the embedding matrix with extra columns, so that the dimension of this matrix is $q \times (d + d_{col})$, where $d_{col}$ is the dimension of the row embedding. @Huang2020 find that including this identifier improves the performance of the model. We compare the results of the TabTransformer model to the other models tested below.

@@

For our experiments, we set up a 5-fold cross validation scheme that is applied to each of the models we introduce. In Sections \ref{models} and \ref{interpretation}, the tables and figures are based on a single fold of the cross validation split. In Section \ref{training-inference}, we provide the cross-validated performance metrics of each of the models.

\textcolor{blue}{Whereas the main focus of our contribution} is on applications of categorical embeddings in predictive modeling, \textcolor{blue}{we have fine-tuned the neural network models in this (and the next) sections using the same 5-fold cross validation scheme described above, but applied to a separate 100k data point subsample of the FEMA data. The results of this fine-tuning are reported in detail in the appendix. The performance metrics in this, and the next, section are based on the optimal models found through this process. Our aim is to show that performant predictive models utilizing categorical embeddings can be developed quite easily, and that these models can be enhanced using suitable architectures utilizing attention mechanisms. As we describe the modeling procedure and results, we provide further commentary on recommended practices in practical applications, especially with respect to the neural network models.}

@@

In Figure \@ref(fig:flood-zone-pca), we plot the learned embeddings for `flood_zone` in Model 3 on the coordinate system defined by the first two principle components. In our case, however, we are not able to identify discernable clusters in the plot. From inspecting the matrix decomposition results, we see that the cumulative proportion of variance for the first two principal components are just \textcolor{blue}{0.38, which is a relatively low proportion,} suggesting that perhaps PCA is not the most appropriate dimensionality reduction technique in this case.

@@

For the GLMs, Models 1 and 3, we follow the standard optimization procedure using iteratively reweighted least squares (IWLS) without regularization. For the neural networks, Models 2 and 4, we utilize the \texttt{Adam} optimizer [@Kingma2014] with a learning rate of \textcolor{blue}{0.001 and a minibatch size of 1024. We split the training set into analysis and assessment sets, containing} 80% and 20% \textcolor{blue}{of the rows, respectively, and further select a} 5% \textcolor{blue}{sample of the training set as a validation set to estimate the out of sample performance of each model. We train on the analysis set for a maximum of 150 epochs, and select the best model from the 150 epochs as measured by the validation. Finally, as mentioned above, in addition to this form of early stopping, we also apply the sropout technique with a rate of dropout of} 2.5\%.

@@

>>>>

@Huang2020 refines this approach in a model they call the TabTransformer, which implements two main changes to the approach shown in Equations \ref{eq:embed_to_attn_to_nn}. The first of these is that, instead of using the matrix of embeddings after processing with self-attention ($X'$) directly in the model, the output of a Transformer model is used instead. The second of these refinements is that for each of the $q$ covariates with an embedding in the matrix $X'$, we include a new row embedding which denotes which covariate each embedding relates to. Thus, in the TabTransformer model, we augment the embedding matrix with extra columns, so that the dimension of this matrix is $q \times (d + d_{col})$, where $d_{col}$ is the dimension of the row embedding. @Huang2020 find that including this identifier improves the performance of the model. We compare the results of the TabTransformer model to the other models tested below. \textcolor{blue}{We refer the reader to Section \ref{attention_models} below where the TabTransformer model is shown diagramatically.}

@@

For our experiments, we set up a 5-fold cross-validation scheme that is applied to each of the models we introduce. In Sections \ref{models} and \ref{interpretation}, the tables and figures are based on a single fold of the cross validation split. In Section \ref{training-inference}, we provide the cross-validated performance metrics of each of the models.

\textcolor{blue}{Whereas the main focus of our contribution} is on applications of categorical embeddings in predictive modeling, \textcolor{blue}{we have fine-tuned the neural network models in this (and the next) sections using the same 5-fold cross validation scheme described above, but applied to a separate 100k data point subsample of the NFIP data. The results of this fine-tuning are reported in detail in the appendix and this separate subsample of the NFIP data is not used for any other purpose in this study. The performance metrics in this, and the next, section are based on the optimal models found through this process. Our aim is to show that performant predictive models utilizing categorical embeddings can be developed quite easily, and that these models can be enhanced using suitable architectures utilizing attention mechanisms. As we describe the modeling procedure and results, we provide further commentary on recommended practices in practical applications, especially with respect to the neural network models.}

@@

In Figure \@ref(fig:flood-zone-pca), we plot the learned embeddings for `flood_zone` in Model 3 on the coordinate system defined by the first two principle components. In our case, however, we are not able to identify discernable clusters in the plot. From inspecting the matrix decomposition results, we see that the cumulative proportion of variance for the first two principal components are just \textcolor{blue}{0.38, which is a relatively low proportion,} suggesting that perhaps PCA is not the most appropriate dimensionality reduction technique in this case.  \textcolor{blue}{Additionally, this indicates that each of the dimensions of the calibrated embeddings are important, implying that the embedding has been well calibrated.}

@@

For the GLMs, Models 1 and 3, we follow the standard optimization procedure using iteratively reweighted least squares (IWLS) without regularization. For the neural networks, Models 2 and 4, we utilize the \texttt{Adam} optimizer [@Kingma2014] with a learning rate of \textcolor{blue}{0.001 and a minibatch size of 1024. Using a five-fold cross-validation approach, we split the dateset into analysis and assessment sets, containing} 80% and 20% \textcolor{blue}{of the rows, respectively, and further select a} 5% \textcolor{blue}{sample of the analysis set as a validation set to estimate the out of sample performance of each model. This is repeated five times on non-overlapping subsets of the dataset. For each of these five splits of the data, we train on the analysis set for a maximum of 150 epochs, and select the best model from the 150 epochs as measured by the validation set. Finally, as mentioned above, in addition to this form of early stopping, we also apply the dropout technique with a rate of dropout of} 2.5\%.

@@

Acknowledgments {.unnumbered}
==========

We thank the two anonymous reviewers for their helpful comments which significantly improved the manuscript.
